Spatiotemporal Multi-Resolution Approximation of the Amari Type Neural Field Model

### 2012-05-21 Multiresolution_framework/ltx/manuscript.tex
minor edits

\begin{abstract}
%\boldmath
\vik{Neural fields are spatially continuous state variables described by integro-differential equations, which are well suited to describe the spatiotemporal evolution of cortical activations on multiple scales. Here we} develop a multi-resolution approximation (MRA) framework for the integro-difference equation (IDE) neural field model based on semi-orthogonal cardinal B-spline wavelets. State and parameter estimation is performed using the Expectation Maximization (EM) algorithm. A synthetic example is provided
to demonstrate the framework.
\end{abstract} 
\begin{keyword}
  Neural Field Model, multi-resolution approximation (MRA), Expectation Maximization (EM) algorithm
\end{keyword}

@@

\section{Multi-Resolution Cortical Architecture}
The human neocortex has an inherent multi-resolution architecture. The spatiotemporal neural field can be written as 
\begin{equation}
	v\left(\mathbf{r},t\right) = \sum_{j} v_j\left(\mathbf{r},t\right),
\end{equation}
where $v\left(\mathbf{r},t\right)$ is a voltage field that may be measured with a multi-electrode array or voltage sensitive dye, $\mathbf{r}\in \mathbb{R}^n$ (where $n\le3$) describes the spatial location within the field, $t$ denotes time, and $j$ indexes the different levels of resolution in the field. This multi-resolution structure is depicted in Fig.~\ref{fig:MultiLayerFieldModel}.
\begin{figure}[t]
	% \centering
		\includegraphics[scale=1]{./Graph/fig1.pdf}
	\caption{{\bf Multi-layer neural field model}. The inset figures show the shape of the
connectivity kernel components at each spatial resolution.}
	\label{fig:MultiLayerFieldModel}
\end{figure}

@@

In this paper, we have presented a novel model-based framework for estimating cortical dynamics from electrophysiological measurements. The novel and key developments of the paper include the multi-resolution finite element model representation of the neural field, and the estimator for an intracortical connectivity kernel with an arbitrary shape. This work is significant, as the ability to create patient-specific neural field models has the potential to contribute to our understanding and improve treatment of diseases resulting from abnormal neurodynamics. Other groups have previously highlighted the importance of a multi-resolution approach in neural field modeling, where the dynamics and connectivity structure differs at different spatial scales~\citep{Jirsa2002,Jirsa2004,Qubbaj2009,Breakspear2005,Schultze-Kraft2011}. 

@@

Additional extension for future work include extending the model to capture a second order synaptic response kernel, The second-order postsynaptic kernel proposed in \citet{VanRotterdam1982} leads to dynamics that are more reminiscent to real postsynaptic potentials. Another extension of this framework is the inclusion of finite distant-dependent action potential propagation velocity into the parametric neural field equations. \vik{In small networks the resulting delays via propagation can be neglected. However, delays must be accounted for in large scale networks. For instance, \citet{Ghosh2008} demonstrated that inclusion of time delays arising from propagation along connecting fibres are essential for the understanding of the fluctuations in resting state networks (see \citet{Deco2010} for review). In fact, \citet{Ghosh2008} emphasize that the couplings in large-scale brain networks have a space-time structure composed of their anatomical connectivity (space) and signal transmission delays (time). To capture correctly the constraints posed upon the network dynamics by connectivity, the space-time structure of the brain connectivity must be correctly estimated.} Finally, future work should be directed towards applying and validating the framework on real data.

>>>>

\begin{abstract}
%\boldmath
\vik{Neural fields are spatially continuous state variables described by integro-differential equations, which are well suited to describe the spatiotemporal evolution of cortical activations on multiple scales. Here we} develop a multi-resolution approximation (MRA) framework for the integro-difference equation (IDE) neural field model based on semi-orthogonal cardinal B-spline wavelets. In this way, a flexible framework is created, whereby both macroscopic and microscopic behavior of the system can be represented simultaneously. State and parameter estimation is performed using the Expectation Maximization (EM) algorithm. A synthetic example is provided
to demonstrate the framework.
\end{abstract} 
\begin{keyword}
  Neural Field Model, multi-resolution approximation (MRA), Expectation Maximization (EM) algorithm
\end{keyword}

@@

\section{Multi-Resolution Cortical Architecture}
The human neocortex has an inherent multi-resolution architecture. The spatiotemporal neural field can be written as 
\begin{equation}
	v\left(\mathbf{r},t\right) = \sum_{j} v_j\left(\mathbf{r},t\right),
\end{equation}
where $v\left(\mathbf{r},t\right)$ is a voltage field that may be measured with a multi-electrode array or voltage sensitive dye, $\mathbf{r}\in \mathbb{R}^n$ (where $n\le3$) describes the spatial location within the field, $t$ denotes time, and $j$ indexes the different levels of resolution in the field. This multi-resolution structure is depicted in Fig.~\ref{fig:MultiLayerFieldModel}.
\begin{figure}[t]
	% \centering
		\includegraphics[scale=1]{./Graph/fig1.pdf}
	\caption{{\bf Multi-layer neural field structure}. The inset figures show the shape of the
connectivity kernel components at each spatial resolution.}
	\label{fig:MultiLayerFieldModel}
\end{figure}

@@

In this paper, we have presented a novel model-based framework for estimating cortical dynamics from electrophysiological measurements. The novel and key developments of the paper include the multi-resolution finite element model representation of the neural field, and the estimator for an intracortical connectivity kernel with an arbitrary shape. This work is significant, as the ability to create patient-specific neural field models has the potential to contribute to our understanding and improve treatment of diseases resulting from abnormal neurodynamics. Other groups have previously highlighted the importance of a multi-resolution approach in neural field modeling, where the dynamics and connectivity structure differs at different spatial scales~\citep{Jirsa2002,Jirsa2004,Breakspear2005,Qubbaj2009,Schultze-Kraft2011}. 

@@

Additional extension for future work include extending the model to capture a second order synaptic response kernel. The second-order postsynaptic kernel proposed in \citet{VanRotterdam1982} leads to dynamics that are more reminiscent to real postsynaptic potentials. Another extension of this framework is the inclusion of finite distant-dependent action potential propagation velocity into the parametric neural field equations. \vik{In small networks the resulting delays via propagation can be neglected. However, delays must be accounted for in large scale networks. For instance, \citet{Ghosh2008} demonstrated that inclusion of time delays arising from propagation along connecting fibres are essential for the understanding of the fluctuations in resting state networks (see \citet{Deco2010} for review). In fact, \citet{Ghosh2008} emphasize that the couplings in large-scale brain networks have a space-time structure composed of their anatomical connectivity (space) and signal transmission delays (time). To capture correctly the constraints posed upon the network dynamics by connectivity, the space-time structure of the brain connectivity must be correctly estimated.} Finally, future work should be directed towards applying and validating the framework on real data.
